School closures and consolidations: the active travel and emission implications of reduced accessibility

### 2024-04-29 paper/Manuscript.qmd
Draft up until Results done

@@

catchment_schools_plot <-
  tm_shape(ALL_SCHOOLS_CATCH_Elem_extra) +
  tm_polygons(col = "OTGC", border.alpha=0.1, palette = "Greys",
              labels = c("[181-318)", "[318-406)", "[406-566)","[566-1417]"),
              breaks = c(181, 318, 406, 566, 1417),
              title = "On the ground\ncapacity (OTGC)")+
  tm_facets(by=c("System","Year"))+
  tm_shape(ALL_SCHOOLS_Elem_formapping) +
  tm_dots(col = "Status", size=0.1, title = "School status between\n2011 and 2016",
          palette = c("#ffd14d", "black", "darkgreen","#0E86D4","#90EE90"),
          labels = c("No change","Closed before 2016", "Expanded before 2016","Moved before 2016","New after 2011")          
  ) +
  tm_facets(by=c("Year","System"))+
  tmap_defaults +
  tm_layout(panel.labels = list(c('HWDCSB (Catholic Public)','HWDSB (Public)'),c('2011','2016')))

tmap_save(catchment_schools_plot, paste0(here::here(),"/Figures/catchment_schools_plot.png"), height=7, width=12, dpi=600)
```

@@

```{r}
SAvail_im_DA2011 <- avail_2011 %>%
  group_by(GeoUID, 
           mode) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT)) |>
  filter(GeoUID != 35250242 & GeoUID != 35250133 & GeoUID != 35250457 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

SAvail_im_DA2016 <- avail_2016 %>%
  group_by(GeoUID, 
           mode) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj = ifelse(is.na(v_im_kid_popadj) | is.nan(v_im_kid_popadj), 0, v_im_kid_popadj),
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT)) |>
  filter(GeoUID != 35250398 & GeoUID != 35250133 & GeoUID != 35250427 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

@@

SAvail_im_DA2016 <- SAvail_im_DA2016 |>
  left_join(cen_data_2016_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> st_set_geometry("geometry")
```
```{r}
#Create 1 sf with data joined, no changes
SAvail_im_DA <-
  rbind(SAvail_im_DA2011 |> mutate(year ="2011"),
        SAvail_im_DA2016 |> mutate(year ="2016")) |>
  mutate(mode = case_when(mode == "c" ~ "motorized",
                          mode == "w" ~ "non-motorized"))

#calc changes between the years - make that 1 sf
SAvail_im_DA_changes <- full_join(SAvail_im_DA2011|>st_drop_geometry(),
                                  SAvail_im_DA2016, by=c("GeoUID", "mode")) |>
  transmute(GeoUID,
            mode = case_when(mode == "c" ~ "motorized",
                             mode == "w" ~ "non-motorized"),
            COMMUNITY = dplyr::coalesce(COMMUNITY.x, COMMUNITY.y),
            c_total_tt = total_tt.y - total_tt.x, #2016 minus 2011 tt
            c_V_im = V_im.y -V_im.x,
            c_V_im_perc = (V_im.y -V_im.x)/V_im.x,
            c_kid_popadj_m = kid_popadj_m.y - kid_popadj_m.x,
            c_DA_pop_5to14 = DA_pop_5to14.y - DA_pop_5to14.x,
            c_DA_prev_LIMAT_under18 = DA_prev_LIMAT_under18.y - DA_prev_LIMAT_under18.x,
            c_emissions_total_car_tt = emissions_total_car_tt.y - emissions_total_car_tt.x,
            c_v_im_kid_popadj = v_im_kid_popadj.y - v_im_kid_popadj.x,
            c_v_im_kid_popadj_perc = (v_im_kid_popadj.y - v_im_kid_popadj.x)/v_im_kid_popadj.x,
            c_v_im_kid_popadj_LIMAT = v_im_kid_popadj_LIMAT.y - v_im_kid_popadj_LIMAT.x,
            c_V_i_kid_popadj_LIMAT = V_i_kid_popadj_LIMAT.y - V_i_kid_popadj_LIMAT.x,
            geometry) |> st_as_sf()

@@

```{r}
#repeat but for all 1 mode
SAvail_i_DA2011 <- avail_2011 %>%
  group_by(GeoUID) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj, na.rm = T),
            V_i = sum(V_ij, na.rm = T),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            kid_pop_DAavg = mean(kid_pop, na.rm = T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm = T),
            .groups = "drop") |>
  filter(GeoUID != 35250242 & GeoUID != 35250133 & GeoUID != 35250457 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

SAvail_i_DA2016 <- avail_2016 %>%
  group_by(GeoUID) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_i = sum(V_ij, na.rm = T),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            kid_pop_DAavg = mean(kid_pop, na.rm = T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm = T),
            .groups = "drop") |>
  filter(GeoUID != 35250398 & GeoUID != 35250133 & GeoUID != 35250427 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

#add the geometries -- census DAs
SAvail_i_DA2011 <- SAvail_i_DA2011 |>
  left_join(cen_data_2011_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> mutate(Year = "2011") |> st_set_geometry("geometry")

SAvail_i_DA2016 <- SAvail_i_DA2016 |>
  left_join(cen_data_2016_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> mutate(Year = "2016") |> st_set_geometry("geometry")

#join them together!
SAvail_i_DA_extra <- rbind(SAvail_i_DA2011, SAvail_i_DA2016)
rm(SAvail_i_DA2011,SAvail_i_DA2016)
```

```{r}
#now do this at the community level
ham_community_bounds <- ham_community_bounds |> 
  mutate(COMMUNITY = case_when(COMMUNITY_ == "Ancaster" ~ "AN",
                               COMMUNITY_ == "Dundas" ~ "DU",
                               COMMUNITY_ == "Flamborough" ~ "FL",
                               COMMUNITY_ == "Glanbrook" ~ "GL",
                               COMMUNITY_ == "Hamilton" ~ "HA",
                               COMMUNITY_ == "Stoney Creek" ~ "SC"))

SAvail_im_COMM2011 <- SAvail_im_DA2011 %>% st_drop_geometry() |>
  group_by(COMMUNITY, 
           mode) %>%
  summarize(total_tt = sum(total_tt, na.rm=T),
            V_im = sum(V_im, na.rm=T),
            kid_popadj_m = sum(kid_popadj_m, na.rm=T),
            COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            emissions_total_car_tt = sum(emissions_total_car_tt, na.rm=T),
            .groups = "drop") |>
  mutate(v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(COMM_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(COMM_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT))  |> left_join(ham_community_bounds |> dplyr::select(-c("OBJECTID")), by=c("COMMUNITY")) |> st_set_geometry("geometry")

SAvail_im_COMM2016 <- SAvail_im_DA2016 %>% st_drop_geometry() |>
  group_by(COMMUNITY, 
           mode) %>%
  summarize(total_tt = sum(total_tt, na.rm=T),
            V_im = sum(V_im, na.rm=T),
            kid_popadj_m = sum(kid_popadj_m, na.rm=T),
            COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            emissions_total_car_tt = sum(emissions_total_car_tt, na.rm=T),
            .groups = "drop") |>
  mutate(v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(COMM_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_im_kid_popadj_LIMAT = V_im*(COMM_prev_LIMAT_under18/100),
         V_im_kid_popadj_LIMAT = ifelse(is.na(V_im_kid_popadj_LIMAT) | is.nan(V_im_kid_popadj_LIMAT), 0, V_im_kid_popadj_LIMAT))  |> left_join(ham_community_bounds |> dplyr::select(-c("OBJECTID")), by=c("COMMUNITY")) |> st_set_geometry("geometry")
```

```{r}
#repeat above but with just i, this is to get the correct LIM value.
SAvail_i_COMM2011 <- SAvail_im_DA2011 %>% st_drop_geometry() |>
  group_by(COMMUNITY) %>%
  summarize(COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            .groups = "drop")

SAvail_i_COMM2016 <- SAvail_im_DA2016 %>% st_drop_geometry() |>
  group_by(COMMUNITY) %>%
  summarize(COMM_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18, na.rm=T),
            .groups = "drop")

#calc changes between the years - make that 1 sf
SAvail_i_COMM_changes <- full_join(SAvail_i_COMM2011 |> st_drop_geometry(),
                                   SAvail_i_COMM2016, by=c("COMMUNITY")) |>
  transmute(COMMUNITY,
            prev_LIMAT_2011 = COMM_prev_LIMAT_under18.x,
            prev_LIMAT_2016 = COMM_prev_LIMAT_under18.y,
            c_prev_LIMAT_perc = ((COMM_prev_LIMAT_under18.y - COMM_prev_LIMAT_under18.x)/
                                   COMM_prev_LIMAT_under18.x )|> scales::percent())
```

>>>>

```{r, include=FALSE}
#boundaries for mapping
hamilton_communities <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/Community_Boundaries.shp")) |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))

ham_bound <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/City_Boundary.shp"))

hydro_p_LakeOntario <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/hydro_p_LakeOntario.shp")) |> st_transform(crs=4326)

ham_bay <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/Waterbodies.shp")) |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")

ham_streets <- st_read(paste0(here::here(),"/data-prep/data-inputs/Boundaries/Street_Centreline.shp")) |> st_transform(crs=4326) |> filter(ROAD_CLASS %in% c("Provincial Highway", "Parkway", "Arterial")) #no collectors, arterial, Local
```

@@

catchment_schools_plot <-
   tm_shape(ham_bay, bbox=hamilton_communities) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=hamilton_communities) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(ALL_SCHOOLS_CATCH_Elem_extra) +
  tm_polygons(col = "OTGC", border.alpha=0.1, palette = "RdPu",
              labels = c("[181-318)", "[318-406)", "[406-566)","[566-1417]"),
              breaks = c(181, 318, 406, 566, 1417),
              title = "On the ground\ncapacity (OTGC)")+
  tm_facets(by=c("System","Year"))+
  tm_shape(ALL_SCHOOLS_Elem_formapping) +
  tm_symbols(shape = "Status",
             size=0.15,
          title.shape = "School status",
          shapes = c(24,25,22,21,23),
          shapes.labels = c(
          "Expanded after 2011",
          "Moved before 2016",
          "New after 2011",
          "No change btwn 2011-16",
          "Closed before 2016")
  ) +
  tm_facets(by=c("System","Year"))+
  tm_credits(c("52 schools\n(18,771 OTGC)","49 schools\n(18,039 OTGC)",
                "95 schools\n(41,796 OTGC)","88 schools\n(40,249 OTGC)"),
             position = c("LEFT", "BOTTOM")) +
  tmap_defaults +
  tm_layout(bg.color = "grey",
            panel.labels = list(c('HWDCSB (Catholic Public)','HWDSB (Public)'),
                                c('2011','2016')),
            panel.label.bg.color = "white")

catchment_schools_plot
```

```{r descriptive-catchment-schools-plot2, eval=FALSE}
tmap_save(catchment_schools_plot, paste0(here::here(),"/Figures/catchment_schools_plot.png"), height=7, width=12, dpi=600)
```

@@

```{r}
#calc spatial availability and other associated variables at the DA 2011 level
SAvail_im_DA2011 <- avail_2011 %>%
  group_by(GeoUID, 
           mode) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT)) |>
  filter(GeoUID != 35250242 & GeoUID != 35250133 & GeoUID != 35250457 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

#calc spatial availability and other associated variables at the DA 2016 level
SAvail_im_DA2016 <- avail_2016 %>%
  group_by(GeoUID, 
           mode) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj = ifelse(is.na(v_im_kid_popadj) | is.nan(v_im_kid_popadj), 0, v_im_kid_popadj),
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT)) |>
  filter(GeoUID != 35250398 & GeoUID != 35250133 & GeoUID != 35250427 ) #these GeoUID's have 0 kid population, so let's drop so we don't have zeros

@@

SAvail_im_DA2016 <- SAvail_im_DA2016 |>
  left_join(cen_data_2016_hamiltononly |> filter(Region.Name == "Hamilton") |> dplyr::select(GeoUID), by="GeoUID") |> st_set_geometry("geometry")

#Create 1 sf with data joined (no changes, just 2011 and 2016)
SAvail_im_DA <-
  rbind(SAvail_im_DA2011 |> mutate(year ="2011"),
        SAvail_im_DA2016 |> mutate(year ="2016")) |>
  mutate(mode = case_when(mode == "c" ~ "motorized",
                          mode == "w" ~ "non-motorized"))
```

```{r}
#calc changes between the years - but at the 2016 DA level 
SAvail_im_DA2011_forGeoUID2016 <- avail_2011 %>%
  group_by(GeoUID_2016, 
           mode) %>%
  summarize(COMMUNITY = first(COMMUNITY),
            total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT))

SAvail_im_DA_changes <- full_join(SAvail_im_DA2011_forGeoUID2016,
                                  SAvail_im_DA2016, by=c("GeoUID_2016"="GeoUID", 
                                                         "mode")) %>%
  transmute(GeoUID_2016,
            mode = case_when(mode == "c" ~ "motorized",
                             mode == "w" ~ "non-motorized"),
            COMMUNITY = dplyr::coalesce(COMMUNITY.x, COMMUNITY.y),
            c_total_tt = total_tt.y - total_tt.x, #2016 minus 2011 tt
            c_total_ttperkid = (total_tt.y/kid_popadj_m.y) - (total_tt.x/kid_popadj_m.x),
            c_V_im = V_im.y -V_im.x,
            c_V_im_perc = (V_im.y -V_im.x)/V_im.x,
            c_kid_popadj_m = kid_popadj_m.y - kid_popadj_m.x,
            c_DA_pop_5to14 = DA_pop_5to14.y - DA_pop_5to14.x,
            c_DA_prev_LIMAT_under18 = DA_prev_LIMAT_under18.y - DA_prev_LIMAT_under18.x,
            c_emissions_total_car_tt = emissions_total_car_tt.y - emissions_total_car_tt.x,
            c_v_im_kid_popadj = v_im_kid_popadj.y - v_im_kid_popadj.x,
            c_v_im_kid_popadj_perc = (v_im_kid_popadj.y - v_im_kid_popadj.x)/v_im_kid_popadj.x,
            c_v_im_kid_popadj_LIMAT = v_im_kid_popadj_LIMAT.y - v_im_kid_popadj_LIMAT.x,
            c_V_i_kid_popadj_LIMAT = V_i_kid_popadj_LIMAT.y - V_i_kid_popadj_LIMAT.x) |>
  left_join(cen_data_2016_hamiltononly|>dplyr::select(c("GeoUID")), 
            by=c("GeoUID_2016" = "GeoUID")) %>% st_as_sf

@@

```{r}
#calc changes between the years - but between communities with geom for mapping
ham_community_bounds <- ham_community_bounds |> 
  mutate(COMMUNITY = case_when(COMMUNITY_ == "Ancaster" ~ "AN",
                               COMMUNITY_ == "Dundas" ~ "DU",
                               COMMUNITY_ == "Flamborough" ~ "FL",
                               COMMUNITY_ == "Glanbrook" ~ "GL",
                               COMMUNITY_ == "Hamilton" ~ "HA",
                               COMMUNITY_ == "Stoney Creek" ~ "SC"))

SAvail_im_COMM2011 <- avail_2011 %>% 
  group_by(COMMUNITY, 
           mode) %>%
  summarize(total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT))

SAvail_im_COMM2016 <- avail_2016 %>%
  group_by(COMMUNITY, 
           mode) %>%
  summarize(total_tt = sum(travel_time_p50*kid_popadj),
            V_im = sum(V_ij),
            kid_popadj_m = sum(kid_popadj),
            DA_pop_5to14 = mean(DA_pop_5to14, na.rm=T),
            DA_prev_LIMAT_under18 = mean(DA_prev_LIMAT_under18),
            .groups = "drop") |>
  mutate(emissions_total_car_tt = ifelse(mode == "c", 0.1*total_tt, 0),
         v_im_kid_popadj = V_im/kid_popadj_m,
         v_im_kid_popadj = ifelse(is.na(v_im_kid_popadj) | is.nan(v_im_kid_popadj), 0, v_im_kid_popadj),
         v_im_kid_popadj_LIMAT = v_im_kid_popadj*(DA_prev_LIMAT_under18/100),
         v_im_kid_popadj_LIMAT = ifelse(is.na(v_im_kid_popadj_LIMAT) | is.nan(v_im_kid_popadj_LIMAT), 0, v_im_kid_popadj_LIMAT),
         V_i_kid_popadj_LIMAT = V_im*(DA_prev_LIMAT_under18/100),
         V_i_kid_popadj_LIMAT = ifelse(is.na(V_i_kid_popadj_LIMAT) | is.nan(V_i_kid_popadj_LIMAT), 0, V_i_kid_popadj_LIMAT))

#add the geometries -- census DAs
SAvail_im_COMM2011 <- SAvail_im_COMM2011 |>
  left_join(ham_community_bounds |> dplyr::select(c("COMMUNITY_","COMMUNITY")), by=c("COMMUNITY"))|> st_set_geometry("geometry")

SAvail_im_COMM2016 <- SAvail_im_COMM2016 |>
  left_join(ham_community_bounds |> dplyr::select(c("COMMUNITY_","COMMUNITY")), by=c("COMMUNITY"))|> st_set_geometry("geometry")

#calc changes between the years - make that 1 sf
SAvail_im_COMM_changes <- full_join(SAvail_im_COMM2011|>st_drop_geometry(),
                                  SAvail_im_COMM2016, by=c("COMMUNITY", "mode")) |>
  transmute(COMMUNITY,
            mode = case_when(mode == "c" ~ "motorized",
                             mode == "w" ~ "non-motorized"),
            c_total_tt = total_tt.y - total_tt.x, #2016 minus 2011 tt
            c_total_ttperkid = (total_tt.y/kid_popadj_m.y) - (total_tt.x/kid_popadj_m.x),
            c_V_im = V_im.y -V_im.x,
            c_V_im_perc = (V_im.y -V_im.x)/V_im.x,
            c_kid_popadj_m = kid_popadj_m.y - kid_popadj_m.x,
            c_DA_pop_5to14 = DA_pop_5to14.y - DA_pop_5to14.x,
            c_DA_prev_LIMAT_under18 = DA_prev_LIMAT_under18.y - DA_prev_LIMAT_under18.x,
            c_emissions_total_car_tt = emissions_total_car_tt.y - emissions_total_car_tt.x,
            c_v_im_kid_popadj = v_im_kid_popadj.y - v_im_kid_popadj.x,
            c_v_im_kid_popadj_perc = (v_im_kid_popadj.y - v_im_kid_popadj.x)/v_im_kid_popadj.x,
            c_v_im_kid_popadj_LIMAT = v_im_kid_popadj_LIMAT.y - v_im_kid_popadj_LIMAT.x,
            c_V_i_kid_popadj_LIMAT = V_i_kid_popadj_LIMAT.y - V_i_kid_popadj_LIMAT.x,
            geometry,
            # kid_2011 = kid_popadj_m.x,
            # kid_2016 = kid_popadj_m.y,
            # c_kid_perc = ((kid_popadj_m.y - kid_popadj_m.x)/kid_popadj_m.x) |> scales::percent(),
            # kid_proportion_2011 = (kid_2011/sum(kid_2011)),
            # kid_proportion_2016 = (kid_2016/sum(kid_2016)),
            # c_kid_proportion_perc = ((kid_proportion_2016 - kid_proportion_2011)/kid_proportion_2011)
            # %>% scales::percent(),
            # kid_proportion_2011_perc = (kid_proportion_2011) |> scales::percent(),
            # kid_proportion_2016_perc = (kid_proportion_2016) |> scales::percent(),
            # V_im_2011 = V_im.x,
            # V_im_2016 = V_im.y,
            # c_V_im_perc = ((V_im.y -V_im.x)/V_im.x) |> scales::percent(),
            # V_im_proportion_2011 = (V_im.x/sum(V_im.x)),
            # V_im_proportion_2016 = (V_im.y/sum(V_im.y)),
            # c_V_im_proportion_perc = ((V_im_proportion_2016 - V_im_proportion_2011)/V_im_proportion_2011)
            # %>% scales::percent(),
            # V_im_proportion_2011_perc = (V_im_proportion_2011) |> scales::percent(),
            # V_im_proportion_2016_perc = (V_im_proportion_2016) |> scales::percent(),
            # v_im_2011 = v_im_kid_popadj.x,
            # v_im_2016 = v_im_kid_popadj.y,
            # c_v_im_perc = ((v_im_kid_popadj.y - v_im_kid_popadj.x)/v_im_kid_popadj.x) |> scales::percent(),
            # total_tt_2011 = total_tt.x,
            # total_tt_2016 = total_tt.y,
            # total_tt_proportion_2011 = (total_tt.x/sum(total_tt.x)),
            # total_tt_proportion_2016 = (total_tt.y/sum(total_tt.y)),
            # c_total_tt_proportion_perc = ((total_tt_proportion_2016 - total_tt_proportion_2011)/total_tt_proportion_2011)
            # %>% scales::percent(),
            # total_tt_proportion_2011 = (total_tt_proportion_2011) |> scales::percent(),
            # total_tt_proportion_2016 = (total_tt_proportion_2016) |> scales::percent(),
            # c_total_tt_perc = ((total_tt.y - total_tt.x)/
            #                      total_tt.x)|> scales::percent(),
            # total_tt_perkid_2011 = total_tt.x/kid_2011,
            # total_tt_perkid_2016 = total_tt.y/kid_2016,
            # c_total_ttperkid_perc = ((total_tt_perkid_2016 - total_tt_perkid_2011)/
            #                            total_tt_perkid_2011)|> scales::percent(),
            # emissions_2011 = emissions_total_car_tt.x,
            # emissions_2016 = emissions_total_car_tt.y,
            # c_emissions_perc = ((emissions_total_car_tt.y - emissions_total_car_tt.x)/
            #                       emissions_total_car_tt.x)|> scales::percent(),
            # emissions_proportion_2011 = (emissions_2011/sum(emissions_2011)),
            # emissions_proportion_2016 = (emissions_2016/sum(emissions_2016)),
            # c_emissions_proportion = ((emissions_proportion_2016 - emissions_proportion_2011)/emissions_proportion_2011)
            # %>% scales::percent(),
            # emissions_proportion_2011_perc = (emissions_proportion_2011) |> scales::percent(),
            # emissions_proportion_2016_perc = (emissions_proportion_2016) |> scales::percent(),
            # emissions_perkid_2011 = emissions_total_car_tt.x/kid_popadj_m.x,
            # emissions_perkid_2016 = emissions_total_car_tt.y/kid_popadj_m.y,
            # c_emissions_perkid_perc = ((emissions_perkid_2016 - emissions_perkid_2011)/
            #                              emissions_perkid_2011)|> scales::percent()
            ) |> st_as_sf()
